The Key to a Performant Data Parallel Compiler

### 2016-11-16 2017-spring-pldi/hsu-2017-key_data_parallel_compiler.tex
Republication Policy conformance

In the following sections we describe the core methods and idioms surrounding the AST encoding, construction of the Path Matrix, and the use of these structures in handling nested, recursive AST relationships. This enables us to perform arbitrary computation over arbitrary sub-trees of an AST selected by parent-child relationships in the manner often seen in compiler transformations. These ideas have been further implemented and tested through the implementation of a complete data-parallel by construction compiler, called Co-dfns, that compiles a lexically scoped, functionally oriented dialect of APL with nested functions \cite{hsu2014co,hsu2015accelerating}. The compiler targets both the GPU and the CPU, and its core is implemented in this pure, restricted language that uses only function composition over data-parallel primitives.

To illustrate the performance of these techniques, we micro-benchmark a transformation heavy expression lifting algorithm that is found almost universally in compilers. This benchmark shows performance speedups of up to a factor of 58 when compared against pure, functionally oriented algorithms.

@@

\item We describe a method of computing over arbitrary sub-trees selected by their parent-child relationships in a data-parallel manner using the Key operator and a Path Matrix.

>>>>

In the following sections we extend, expound, and clarify on existing work \cite{hsu2016key} to describe the core methods and idioms surrounding the AST encoding, construction of the Path Matrix, and the use of these structures in handling nested, recursive AST relationships, which enable us to perform arbitrary computation over arbitrary sub-trees of an AST selected by parent-child relationships in the manner often seen in compiler transformations. These ideas have been further implemented and tested through the implementation of a complete data-parallel by construction compiler, called Co-dfns, that compiles a lexically scoped, functionally oriented dialect of APL with nested functions \cite{hsu2014co,hsu2015accelerating}. The compiler targets both the GPU and the CPU, and its core is implemented in this pure, restricted language that uses only function composition over data-parallel primitives. To illustrate the performance of these techniques, we benchmark a transformation heavy expression lifting algorithm that is found almost universally in compilers. This benchmark shows performance speedups of up to a factor of 58 when compared against pure, functionally oriented algorithms.

@@

\item We expound on a method \cite{hsu2016key} of computing over arbitrary sub-trees selected by their parent-child relationships in a data-parallel manner using the Key operator and a Path Matrix.
